Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2020-12-23 GWLDCII.tex
more from the phone conversation

Wilson loop diagrams, as we will study them, are objects given by a perturbative expansion of the holomorphic Wilson loop in twistor space.
Classically, the interactions of the theory were studied as maximal helicity violating diagrams ($MHV$ diagrams), next-to-maximal helicity violating diagrams and so on ($N^kMHV$ diagrams) in momentum space. Adamo and Mason, in \cite{Adamo:2011pr}, give an elegant formulation for the $N^kMHV$ diagrams in twistor space. The Wilson loop diagram perspective of this paper is dual to this twistor representation of $MHV$ diagrams.
The objects are closely related and carry the same information.  We choose to emphasize the useful dual perspective by our focus on the Wilson loop diagrams.

@@

In \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between theWilson loop diagrams and positroids, defined as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure zero \cite{WLDdim}.

In the above work and elsewhere, it is clear that while the space parametrized by the Wilson loop diagrams is related to the Amplituhedron, they are not the same space, nor is a direct connection between the geometry of the Amplituhedron and the geometry of the Wilson loop diagrams known. For instance, it is conjectured that the Amplituhedron is orientable, whereas Agarwala and Marcott show in \cite{non-orientable} that the space parametrized by the Wilson loop diagrams as a subspace of $\Gr(k, n+1)$ can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, which can be seen by comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. Furthermore, the spaces in the two situations have different dimensions, as we will discuss further below, and examples show\cite{HeslopSteward} that they do not relate by a coordinate projection or boundary map or other straightforward function.  None the less, from their common physical origin, it is believed that once the integrals over these spaces are taken the resulting amplitudes will show a more direct relation \cite{wilsonloop, Amplituhedronsquared}.

@@

This paper continues the program of establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. In \cite{non-orientable} the authors note that the Wilson loop diagrams only correspond to positroid cells when the gauge vector is ignored. Including the gauge vector defines a $4k$ dimensional subspace of $\mathbb{G}r_\R(k,n+1)$.\st{This apparent contradiction}\todo{I think we should say more in view of reviewer Y's comment 3 and we can probably phrase it more positively too} \st{is resolved by the non-orientability result of {\cite{non-orientable}}.}

As described above, each integrand associated to a Wilson loop diagram gives the volume of a convex polytope. However, each individual integral has singularities which come from the factors in the denominator of the integrands. SYM $N=4$ is a finite theory. However at tree level, the integrals used to calculate the amplitudes of the theory have singularities upon evaluation \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}. These singularities are classified as spurious and physical in the literature, and the spurious poles must all cancel in the calculation of a tree level scattering amplitude,\sanote{Does this address the reviewer's concern properly?} though how this works out mathematically is unclear. The cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$-dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto \cite{generalcombinatoricsI}.

@@

All three of these conditions arise from the physical interpretation of Wilson loop diagrams.  The first condition comes from the fact that these diagrams represent $N^k$MHV interctions, where $k +4 <n$, see for instance the explanation in the introduction of \cite{adamo:2011pr}.  The second condition can be derived from the integrals associated to the diagrams, as in \cite[Theorem 1.13]{wilsonloop}. Note that the second condition prohibits propagators that connect two adjacent edges. In our definition, $\cP$ is a set so we may not have two propagators starting and ending on the same pair of edges, but we have lost nothing of physical value by not allowing multiple propagators ending on the same pair of edges since these would also fail the second condition. The third condition comes from the fact that these Wilson loop diagrams are a representation of planar Feynman diagrams of $N=4$ SYM theory. For a description of the relationship between planar Feynman diagram and planar Wilson loops, see the discussion in \cite[section 3]{LipsteinMason}.
\hlfix{}{I changed the previous 2 paragraphs quite a bit, to separate the math intuition from the physical reasoning.}  \sanote{fixed cittation issues. Like the  rewrite}

>>>>

Wilson loop diagrams, as we will study them, are objects given by a perturbative expansion of the holomorphic Wilson loop in twistor space.
Classically, the interactions of the theory were studied as maximal helicity violating diagrams ($MHV$ diagrams), next-to-maximal helicity violating diagrams and so on ($N^kMHV$ diagrams) in momentum space. Adamo and Mason, in \cite{Adamo:2011pr}, give an elegant formulation for the $N^kMHV$ diagrams in twistor space. The Wilson loop diagram perspective of this paper is dual to this twistor representation of $MHV$ diagrams.
The objects are closely related and carry the same information.  We choose to emphasize the dual perspective by our focus on the Wilson loop diagrams.

@@

In \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Wilson loop diagrams and positroids, defined as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure zero \cite{WLDdim}.

In the above work and elsewhere, it is clear that while the space parametrized by the Wilson loop diagrams is related to the Amplituhedron, they are not the same space, nor is a direct connection between the geometry of the Amplituhedron and the geometry of the Wilson loop diagrams known. For instance, it is conjectured that the Amplituhedron is orientable, whereas Agarwala and Marcott show in \cite{non-orientable} that the space parametrized by the Wilson loop diagrams as a subspace of $\Gr(k, n+1)$ can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, which can be seen by comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. Furthermore, the spaces in the two situations have different dimensions, as we will discuss further below, and examples show \cite{HeslopStewart} that they do not relate by a coordinate projection or boundary map or other straightforward function.  None the less, from their common physical origin, it is believed that once the integrals over these spaces are taken the results will parametrize related spaces \cite{wilsonloop, Amplituhedronsquared}.

@@

This paper continues the program of establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. In \cite{non-orientable} the authors note that the Wilson loop diagrams only correspond to positroid cells when the gauge vector is ignored. Including the gauge vector defines a $4k$ dimensional subspace of $\mathbb{G}r_\R(k,n+1)$.

As described above, each integrand associated to a Wilson loop diagram gives the volume of a convex polytope. However, each individual tree level integral has singularities. These singularities are of two types physical and spurious.  The physical singularities come from special kinematical points and we will not consider them in this paper.  The spurious singularities come from zeros of the denominator, and hence are accessible to our combinatorial techniques.  SYM $N=4$ is a finite theory so we only expect to see the physical singularites in the overall amplitude.  Consequently, the spurious singularities are artifacts of how the integrals are represented and must cancel in the calculation of a tree level scattering amplitude, though the details of how this works is unclear. The cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$-dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto \cite{generalcombinatoricsI}.

%SYM $N=4$ is a finite theory.  which come from the factors in the denominator of the integrands.  However at tree level, the integrals used to calculate the amplitudes of the theory have singularities upon evaluation \cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}. These singularities are classified as spurious and physical in the literature, and the spurious poles must all cancel in the calculation of a tree level scattering amplitude,\sanote{Does this address the reviewer's concern properly?} though how this works out mathematically is unclear. The cancellation has been conjectured in general \cite{Hodges2013} and verified explicitly in a few cases \cite{casestudy, HeslopStewart}, but it is hard to prove in general due to the fact that the relationship between Wilson loop diagrams and $3k$-dimensional positroid cells of $\Gr(k,n)$ is neither one-to-one nor onto \cite{generalcombinatoricsI}.

@@

All three of these conditions arise from the physical interpretation of Wilson loop diagrams.  The first condition comes from the fact that these diagrams represent $N^k$MHV interctions, where $k +4 <n$, see for instance the explanation in the introduction of \cite{Adamo:2011pr}.  The second condition can be derived from the integrals associated to the diagrams, as in \cite[Theorem 1.13]{wilsonloop}. Note that the second condition prohibits propagators that connect two adjacent edges. In our definition, $\cP$ is a set so we may not have two propagators starting and ending on the same pair of edges, but we have lost nothing of physical value by not allowing multiple propagators ending on the same pair of edges since these would also fail the second condition. The third condition comes from the fact that these Wilson loop diagrams are a representation of planar Feynman diagrams of $N=4$ SYM theory. For a description of the relationship between planar Feynman diagram and planar Wilson loops, see the discussion in \cite[section 3]{LipsteinMason}.
\hlfix{}{I changed the previous 2 paragraphs quite a bit, to separate the math intuition from the physical reasoning.}  \sanote{fixed cittation issues. Like the  rewrite}
